Centralizers of higher enveloping algebras of Lie algebras

### 2020-01-28 draftbygreg.tex
outlining strategy

%Recently, a generalization of the concept of the enveloping algebra of a Lie algebra has emerged: each dg Lie algebra $\fg$ has an enveloping $E_n$ algebra $\UU_n\fg$, for every $n$, so that $\UU_1\fg$ is the usual enveloping algebra $U\fg$. Building on prior work \cite{BD, AF} [Owen's thesis?], Knudsen \cite{Knudsen} developed these ideas extensively and offered a useful factorization algebra model for the enveloping $E_n$ algebra. Our central goal here is to generalize the following results about the enveloping algebra $U\fg$ to $\UU_n\fg$, for all $n$.
%
%\begin{enumerate}
%\item[(1)] the (underived) center of $U\sl_{N+1}$ is isomorphic to $(\Sym(\CC^{N}))^{S_N}$, where $\CC^{N}$ denotes the Cartan subalgebra of $\sl_{N+1}$ and $S_N$ is the Weyl group of $SL_{N+1}$, and
%\item[(2)] the derived center of the enveloping algebra $U\fg$ is given by derived invariants of the adjoint action of $\fg$ on $U\fg$. That is to say $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$.
%\end{enumerate}
%
%Our results pertains to (1) by taking $N \to \infty$, and to (2) through the derived {\em centralizer} of the map $\UU_n\sl_\infty \hookrightarrow \UU_n\gl_\infty(A)$, where $A$ is a dg algebra. Note that the usual derived center is the centralizer of the identity map $\UU_n\fg \to \UU_n\fg$.
%
%We will review these notions and describe our results in detail Min the next section. When $n=1$,  \owen{state the classical thing here!}
%
%Our primary tool here is the Loday-Quillen-Tsygan (LQT) theorem which describes the large $N$ limit of the Lie algebra homology of $\gl_N(A)$, for a dg associative algebra $A$. Since the enveloping $E_n$ algebra of a Lie algebra can be constructed as a locally constant factorization algebra made out of the shifted Chevalley-Eilenberg chains \cite{}, the relevance of the LQT theorem in this context  should be apparent. From this point of view, one can think of our main result as an $E_n$ generalization of the LQT theorem. We hope these ideas have exciting applications in quantum field theory and random matrix theory, as well as to other large $N$ limit phenomena.

@@

\begin{rmk}
We state our theorems in terms of the stable $\infty$-category of cochain complexes 
instead of the stable model categories or pre-triangulated dg categories. 
Many of our constructions, however, involve explicit manipulations of standard complexes, 
and when it is helpful, we indicate the level at which we are working.
\end{rmk}

@@

See section~\ref{sec: enveloping} for a brief discussion of the enveloping $E_n$ algebra construction, and see section~\ref{sec: centralizer} 
for a brief discussion about centralizers to maps of $E_n$ algebras.

\owen{An approximate version of our main theorem. It's not terrible but I'd like to make the statement more accurate and perhaps punchier.}

\begin{thm}
\label{thm: centralizer}
Let $\UU_n \fg$ be the enveloping $E_n$ algebra of a dg Lie algebra $\fg$ and $f: \UU_n \fg \to B$ a map of augmented $E_n$ algebras that gives $B$
the structure of an $E_n$ algebra in $\fg$-modules\footnote{see Proposition~\ref{P:inducedEngModStruct}}. 
The centralizer ${\frak Z}_{E_n}(f)$ is an $E_n$-algebra modeled by the Chevalley-Eilenberg 
cochains $\clies(\fg, B)$ equipped with the convolution $E_n$-product.
\end{thm}

\owen{Here is an outline of the proof strategy, with ingredients indicated:}

Existing work gives us some insight into the situation:
\begin{itemize}
\item Lurie and Francis give a characterization of the centralizer of any map $f: A \to B$ as the ``internal hom'' in left $\int_{S^{n-1}}A$ -modules from $A$ to $B_f$, where $B_f$ denotes $B$ viewed as a left module via the map~$f$.
\item Knudsen verifies that we can model the enveloping $E_n$ algebra by a strict prefactorization algebra $\cliel(\Omega_c \otimes \fg)$ on $\RR^n$, which is weakly locally constant.
\item In combination, we have an explicit characterization of the centralizer up to quasi-isomorphism, 
since it is easy to compute the factorization homology of the enveloping algebra $\UU_n \fg$ over a sphere.
\end{itemize}

Our approach has the following steps:
\begin{enumerate}
\item Transfer the question to the setting of strict prefactorization algebras on $\RR^n$ that are weakly locally constant. (First use Lurie to explain that we can work with the homotopy coherent version, and then invoke strictification theorems from the model category literature on operads.) 
\item In particular, explain why $B$ determines some prefactorization algebra that I'll also call~$B$. Explain how the map $f: \UU_n \fg \to B$ determines a $\fg$-action on $B$. That is, we can obtain a prefactorization algebra $B_f$ taking values in the 1-category of $\fg$-representations.
\item Construct an explicit strict prefactorization algebra $\clies(\Omega \otimes \fg, B_f)$ as a kind of ``convolution algebra.'' Explain why it canonically fits into a centralizer diagram: there is a map
\[
\clies(\Omega \otimes \fg, B_f) \otimes \UU_n \fg \to B_f
\]
such that precomposing with the map $1 \otimes \id_{\UU_n \fg}$ is the map~$f$.
\item Fitting into such a diagram guarantees a canonical map $\clies(\Omega \otimes \fg, B_f) \to \cZ_f$ as strict prefactorization algebras.
\item Verify that this map is a quasi-isomorphism on disks. 
Hence it induces a weak equivalence of the associated $E_n$ algebras. 
\end{enumerate}

Some easy but important situations to examine:
\begin{itemize}
\item When $n=0$, we see naive endomorphisms arise.
\item When $n=1$, we recover standard Koszul duality and Hochschild cohomology examples.
\item Consider $\fg$ {\em abelian} for arbitrary $n$ and compare to the $E_n$ HKR theorem.
\item Compare with $E_n$ HKR theorem for $\clies(\fg)$ via $E_n$ Koszul duality.
\item Use these facts to compute factorization homology of centralizers on manifolds with boundary --- kind of like Chern-Simons states!
\end{itemize}

\owen{End of strategy}

>>>>

Recently, a generalization of the concept of the enveloping algebra of a Lie algebra has emerged: each dg Lie algebra $\fg$ has an enveloping $E_n$ algebra $\UU_n\fg$, for every $n$, so that $\UU_1\fg$ is the usual enveloping algebra $U\fg$. Building on prior work \cite{BD, AF} [Owen's thesis?], Knudsen \cite{Knudsen} developed these ideas extensively and offered a useful factorization algebra model for the enveloping $E_n$ algebra. Our central goal here is to generalize the following results about the enveloping algebra $U\fg$ to $\UU_n\fg$, for all $n$.

\begin{enumerate}
\item[(1)] the (underived) center of $U\sl_{N+1}$ is isomorphic to $(\Sym(\CC^{N}))^{S_N}$, where $\CC^{N}$ denotes the Cartan subalgebra of $\sl_{N+1}$ and $S_N$ is the Weyl group of $SL_{N+1}$, and
\item[(2)] the derived center of the enveloping algebra $U\fg$ is given by derived invariants of the adjoint action of $\fg$ on $U\fg$. That is to say $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$.
\end{enumerate}

Our results pertains to (1) by taking $N \to \infty$, and to (2) through the derived {\em centralizer} of the map $\UU_n\sl_\infty \hookrightarrow \UU_n\gl_\infty(A)$, where $A$ is a dg algebra. Note that the usual derived center is the centralizer of the identity map $\UU_n\fg \to \UU_n\fg$.

We will review these notions and describe our results in detail Min the next section. When $n=1$,  \owen{state the classical thing here!}

Our primary tool here is the Loday-Quillen-Tsygan (LQT) theorem which describes the large $N$ limit of the Lie algebra homology of $\gl_N(A)$, for a dg associative algebra $A$. Since the enveloping $E_n$ algebra of a Lie algebra can be constructed as a locally constant factorization algebra made out of the shifted Chevalley-Eilenberg chains \cite{}, the relevance of the LQT theorem in this context  should be apparent. From this point of view, one can think of our main result as an $E_n$ generalization of the LQT theorem. We hope these ideas have exciting applications in quantum field theory and random matrix theory, as well as to other large $N$ limit phenomena.

@@

\begin{rmk}
We state our theorems in terms of the stable $\infty$-category cochain complexes instead of the appropriate stable model categories or pre-triangulated dg categories. Many of our constructions, however, involve explicit manipulations of standard complexes, and when it is helpful, we indicate the level at which we are working.
\end{rmk}

@@

See section~\ref{sec: enveloping} for a summary of the enveloping $E_n$ algebra construction, and see section~\ref{sec: centralizer} 
for a summary of centralizers of maps of $E_n$ algebras.

\begin{thm}
\label{thm: centralizer}
Let $\UU_n \fg$ be the enveloping $E_n$ algebra of a dg Lie algebra $\fg$ and $f: \UU_n \fg \to B$ a map of augmented $E_n$ algebras which gives $B$
the structure of an $E_n$ algebra in $\fg$-modules\footnote{see Proposition~\ref{P:inducedEngModStruct}}. 
Then, the centralizer ${\frak Z}_{E_n}(f)$ is the $E_n$-algebra given by the Chevalley-Eilenberg 
algebra $\clies(\fg, B)$ with the convolution $E_n$-product.
\end{thm}
